---
title: "Probability"
output: html_document
date: "2022-08-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div class="instructions">
Complete all **Exercises**, and submit answers to **Questions** on the Coursera 
platform.
</div>


## Probability

Basketball players who make several baskets in succession are described as having a hot hand. Fans and players have long believed in the hot hand phenomenon, which refutes the assumption that each shot is independent of the next. However, a 1985 paper by Gilovich, Vallone, and Tversky collected evidence that contradicted this belief and showed that successive shots are independent events. This paper started a great controversy that continues to this day, as you can see by Googling hot hand basketball.

We do not expect to resolve this controversy today. However, in this lab we’ll apply one approach to answering questions like this. The goals for this lab are to (1) think about the effects of independent and dependent events, (2) learn how to simulate shooting streaks in R, and (3) to compare a simulation to actual data in order to determine if the hot hand phenomenon appears to be real.


### Load packages

```{r load-packages, message=FALSE}
library(statsr)
library(dplyr)
library(ggplot2)
```


### Data

```{r load-data}
data(kobe_basket)
```

Counting streak lengths manually for all 133 shots would get tedious, so we'll
use the custom function `calc_streak` to calculate them, and store the results
in a data frame called `kobe_streak` as the `length` variable.

```{r calc-streak-kobe}
kobe_streak <- calc_streak(kobe_basket$shot)
```

We can then take a look at the distribution of these streak lengths.

```{r plot-streak-kobe}
ggplot(data = kobe_streak, aes(x = length)) +
  geom_histogram(binwidth = 1)
```

## Simulations in R

While we don't have any data from a shooter we know to have independent shots, that sort of data is very easy to simulate in R. In a simulation, you set the ground rules of a random process and then the computer uses random numbers to generate an outcome that adheres to those rules. As a simple example, you can simulate flipping a fair coin with the following.

```{r head-tail}
coin_outcomes <- c("heads", "tails")
sample(coin_outcomes, size = 1, replace = TRUE)
```

If you wanted to simulate flipping a fair coin 100 times, you could either run the function 100 times or, more simply, adjust the `size` argument, which governs how many samples to draw (the `replace = TRUE` argument indicates we put
the slip of paper back in the hat before drawing again). Save the resulting vector of heads and tails in a new object called `sim_fair_coin`.

```{r sim-fair-coin}
sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)
```

To view the results of this simulation, type the name of the object and then use `table` to count up the number of heads and tails.

```{r table-sim-fair-coin}
sim_fair_coin
table(sim_fair_coin)
```

Say we're trying to simulate an unfair coin that we know only lands heads 20% of the time. We can adjust for this by adding an argument called `prob`, which provides a vector of two probability weights.

```{r sim-unfair-coin}
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE,prob = c(0.2, 0.8))
```

In a sense, we've shrunken the size of the slip of paper that says "heads", making it less likely to be drawn and we've increased the size of the slip of paper saying "tails", making it more likely to be drawn. When we simulated the 
fair coin, both slips of paper were the same size. This happens by default if you don't provide a `prob` argument; all elements in the `outcomes` vector have an equal probability of being drawn.


## Simulating the Independent Shooter

Simulating a basketball player who has independent shots uses the same mechanism that we use to simulate a coin flip. To simulate a single shot from an independent shooter with a shooting percentage of 50% we type,

```{r sim-basket}
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 1, replace = TRUE)
```


To make a valid comparison between Kobe and our simulated independent shooter, 
we need to align both their shooting percentage and the number of attempted shots.


<div class="exercise">
What change needs to be made to the `sample` function so that it reflects a shooting percentage of 45%? Make this adjustment, then run a simulation to sample 133 shots. Assign the output of this simulation to a new object called `sim_basket`.
</div>
```{r}
# type your code for the Exercise here, and Knit
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE, prob = c(0.45,0.55))
sim_basket
table(sim_basket)
```

   
Note that we've named the new vector `sim_basket`, the same name that we gave to
the previous vector reflecting a shooting percentage of 50%. In this situation, 
R overwrites the old object with the new one, so always make sure that you don't
need the information in an old vector before reassigning its name.

With the results of the simulation saved as `sim_basket`, we have the data 
necessary to compare Kobe to our independent shooter.

Both data sets represent the results of 133 shot attempts, each with the same 
shooting percentage of 45%. We know that our simulated data is from a shooter 
that has independent shots. That is, we know the simulated shooter does not have
a hot hand.

### Comparing Kobe Bryant to the Independent Shooter

<div class="exercise">
Using `calc_streak`, compute the streak lengths of `sim_basket`, and save the results in a data frame called `sim_streak`. Note that since the `sim_streak` object is just a vector and not a variable in a data frame, we don't need to first select it from a data frame like we did earlier when we calculated the streak lengths for Kobe's shots.
</div>
```{r sim-streak-lengths}
# type your code for the Exercise here, and Knit
sim_streak <- calc_streak(sim_basket)

```
   
<div class="exercise">
Make a plot of the distribution of simulated streak lengths of the independent shooter. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? How long is the player's longest streak of baskets in 133 shots?
</div>
```{r plot-sim-streaks}
# type your code for the Exercise here, and Knit
ggplot(data = sim_streak, aes(x=length)) + geom_histogram(binwidth = 1)
```

