---
title: "Week_3 Exam"
output: word_document
date: '2022-07-21'
---

```{r}
library(dplyr)
library(ggplot2)
library(statsr)
```

```{r}
data("arbuthnot")
```


### 1. How many variables are included in this data set (data set: arbuthnot)?

```{r}
names(arbuthnot)
```

### 2 What command would you use to extract just the counts of girls born? 

```{r}
arbuthnot$girls
```

### 3. Which of the following best describes the number of girls baptised over the years included in this dataset? 

```{r}
ggplot(data = arbuthnot, aes(x = year, y = girls)) +
  geom_point()+ geom_line()
```


### 4. How many variables are included in this data set (data set: present)?

```{r}
data("present")
names(present)
```

### 5. Create a new variable called more_boys which contains the value of either TRUE if that year had more boys than girls, or FALSE if that year did not. Based on this variable which of the following statements is true?

```{r}
present <- present %>%
  mutate(more_boys = boys > girls)
  ggplot(data=present, aes(x=year, y=more_boys))+
  geom_point() + geom_line()
```


### 7. Calculate the boy-to-girl ratio each year, and store these values in a new variable called prop_boy_girl in the present dataset. Plot these values over time. Which of the following best describes the trend?


```{r}
present <- present %>%
  mutate(prop_boy_girl = boys/girls)

ggplot(data = present, aes(x=year, y=prop_boy_girl)) +
  geom_point() +geom_line()
```

### 8 In what year did we see the most total number of births in the U.S.? 

```{r}
present <- present %>%
  mutate(total = boys + girls)%>%
  arrange(desc(total))  
```







