---
title: 'EDA: Dr. Arbuthnot’s Baptism Records'
output:
  word_document: default
  html_document: default
date: '2022-07-19'
---

## loading the R Packages

```{r}

library(dplyr)
library(ggplot2)
library(statsr)

```

## Dataset 1: Dr. Arbuthnot’s Baptism Records

The dataset is regarding the arbuthnot baptisum count of boys and girls from 1629 to 
1710. The Arbuthnot data set refers to Dr. John Arbuthnot, an 18th century physician, writer, and mathematician. He was interested in the ratio of newborn boys to newborn girls, so he gathered the baptism records for children born in London for every year from 1629 to 1710. 


```{r}
data("arbuthnot")
head(arbuthnot)
```
## You can see the dimensions of this data frame by typing:

In R row comes first and column later, thus here its indicates there are 82 rows and 3 columns.

```{r}
dim(arbuthnot)
```

## You can see the names of these columns (or variables) by typing:

```{r}
names(arbuthnot)
```
## What years are included in this dataset?

```{r}
arbuthnot$year
```

# Exploratory Data Analysis 

##  number of boys baptized each year

$ --> Using this function we can  access the data in a single column of a data frame separately.

```{r}
arbuthnot$boys
```


## What command would you use to extract just the counts of girls born?


```{r}
arbuthnot$girls
```


## a simple plot of the number of girls baptized per year with the command

Synthax explanation : 

* The first argument is always the dataset.
* aes -->  the x and the y axes.
* + use for for another layer,
```{r}
ggplot(data = arbuthnot, aes(x = year, y = girls)) +
  geom_point()
```

To seek help regarding the ggplot

```{r}
?ggplot
```

## Total number of baptisms 

```{r}
arbuthnot$boys + arbuthnot$girls
```

## Adding a new variable to the data frame

The %>% operator is called the piping operator.Takes the output of the current line and pipes it into the following line of code.

Take the arbuthnot dataset and pipe it into the mutate function. Using this mutate a new variable called total that is the sum of the variables called boys and girls. Then assign this new resulting dataset to the object called arbuthnot, i.e. overwrite the old arbuthnot dataset with the new one containing the new variable.


```{r}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
```

```{r}
ggplot(data = arbuthnot, aes(x=year, y=total)) +
  geom_line()
```

```{r}
ggplot(data = arbuthnot, aes(x = year, y = total)) +
  geom_line() +
  geom_point()
```

## Now, generate a plot of the proportion of boys born over time.

```{r}
ggplot(data = arbuthnot, aes(x = year, y=boys)) +
  geom_line()+
  geom_point()
```

## if boys outnumber girls in each year with the expression ?

```{r}
arbuthnot <- arbuthnot %>%
  mutate(more_boys = boys > girls)
```
